### 2014-11-13 heat-transfer.tex
space saving edits

To further improve locality, we have split the second rule to avoid sending
small $\delta$ values if the target node is in another thread
(Fig.~\ref{code:ht_better}). The \textbf{Local-Only/Regular} line
in Fig.~\ref{results:ht} shows that this has the best performance.  However, this
come at the price of increased errors in the computed heat values.

@@

\begin{topfig}
   \begin{center}
      \includegraphics[width=\figsize]{results/new-heat-transfer-80}
   \end{center}
   \scap{results:ht}{Experimental results for the HT program. The coordinated version
      is, on average, 30\% faster than the regular version although it has a
      slightly worse scalability due to a reduction in work available.}
\end{topfig}

>>>>

To further improve locality, we split the second rule to avoid sending
small $\delta$ values if the target node is in another thread
(Fig.~\ref{code:ht_better}). The \textbf{Local-Only/Regular} line in
Fig.~\ref{results:ht} shows that this performs best.
However, this come at the price of increased errors in the computed
heat values.

@@

\begin{topfig}
   \begin{center}
      \includegraphics[width=\figsize]{results/new-heat-transfer-80}
   \end{center}
   \scap{results:ht}{Experimental results for the HT program. The coordinated version
      is, on average, 30\% faster than the regular version although it has a
      slightly worse scalability due to a reduction in work available.}
\vspace*{-2ex}
\end{topfig}

### 2014-11-13 implementation.tex
space saving edits

The compiler translates each rule to a procedure and a list of facts that need
to exist to satisfy the rule. This procedure is executed by the virtual
machine whenever all the facts mentioned in the body exist.
The procedure loops over all possible combinations of the rule, retrieving facts from the
database, performing join operations and then consuming and deriving facts.
The compiler implements join optimizations (to support efficient rule filtering)
and fact updates (to avoid allocations).

@@

\begin{tightdescription}
   \item[working]: the node is executing.
   \item[inactive]: the node is inactive, i.e., it has no new facts and is not in any
   queue for processing.
   \item[queue]: the node is active with new facts and is in some queue waiting
   to be processed.
   \item[stealing]: the node has just been stolen and is in the process of being
   moved to another thread.
   \item[coordinating]: the node is moving from one queue to another queue.
\end{tightdescription}

@@

In the paper~\cite{cruz-ppdp14}, the LM virtual machine was measured and
compared against implementations of similar programs but written in other
languages. For instance, the Loopy Belief Propagation program was found to be around 2 times slower
than GraphLab, while the N Queens program was found to be around 10 to 15 slower
than a sequential C implementation. Some of those programs were also compared against
a Python implementations and LM faired fairly well. We measured
the impact of the coordination mechanisms implemented for the CLM
virtual machine on regular LM programs and saw very little to no degradation.
Since we also performed general performance improvements, programs such as
PageRank and Greedy Graph Coloring now perform slightly better than before.

>>>>

The compiler translates each rule to a procedure and a list of facts needed
to satisfy the rule. This procedure is executed by the virtual
machine whenever all the facts mentioned in the body exist.
The procedure loops over all possible combinations of the rule, retrieving facts from the
database, performing join operations and then consuming and deriving facts.
The compiler implements join optimizations (to support efficient rule filtering)
and fact updates (to avoid allocations).

@@

\begin{tightdescription}
   \item[working]: the node is executing.
   \item[inactive]: the node is inactive, i.e., it has no new facts and is not in any
   queue for processing.
   \item[queue]: the node is active with new facts and is in some queue waiting
   to be processed.
   \item[stealing]: the node has just been stolen and is in the process of being
   moved to another thread.
   \item[coordinating]: the node is moving from one queue to another.
\end{tightdescription}

@@

In the paper~\cite{cruz-ppdp14}, LM programs were compared to similar
programs written in other languages. The Loopy Belief Propagation
program was around 2 times slower than GraphLab, while the N Queens
program was 10 to 15x slower than a sequential C implementation. The
LM programs had similar performance to programs written in Python.
The introduction of coordination in CLM has little or no impact on the
overall performance of these programs.  In fact, due to some
improvements in the virtual machine, programs such as PageRank and
Greedy Graph Coloring now perform slightly better than before.

### 2014-11-13 splash-bp.tex
space saving edits

\begin{dblfig}
   \begin{center}
      \subfloat[]{\includegraphics[width=\figsize]{results/system_belief-propagation-400.png}}
      \subfloat[]{\includegraphics[width=\figsize]{results/system_splash-bp-400.png}}
      \subfloat[]{\includegraphics[width=\figsize]{results/system_improve_belief-propagation-400.png}}
   \end{center}
   \scap{results:splash_bp}{Experimental results for LBP and SBP. (a) shows the
      scalability of LBP for both CLM and GraphLab and (b) shows the
      scalability of SBP. (c) presents the
      improvements seen in SBP against LBP, where SBP runs, on average, 1.5 to 2.5
      times faster than LBP.}
\end{dblfig}

In this program, coordination assumes a far more important role than
we have seen before. Coordination rules fully drive the behavior of
the algorithm and although the final result of the algorithm is
statistically identical to the original algorithm, SBP works very
differently than LBP.  SBP is also implemented in
GraphLab~\cite{GraphLab2010}, a C++ framework for writing machine
algorithms.  GraphLab provides the splash scheduler as part of its the
framework. It includes 350 lines of C++ code.  With our coordination facts,
it is possible to create the necessary scheduling with only 12 rules.

>>>>

\begin{dblfig}
\vspace*{-1ex}
   \begin{center}
      \subfloat[]{\includegraphics[width=\figsize]{results/system_belief-propagation-400.png}}
      \subfloat[]{\includegraphics[width=\figsize]{results/system_splash-bp-400.png}}
      \subfloat[]{\includegraphics[width=\figsize]{results/system_improve_belief-propagation-400.png}}
   \end{center}
   \scap{results:splash_bp}{Experimental results for LBP and SBP. (a) shows the
      scalability of LBP for both CLM and GraphLab and (b) shows the
      scalability of SBP. (c) presents the
      improvements seen in SBP against LBP, where SBP runs, on average, 1.5 to 2.5
      times faster than LBP.}
\vspace*{-1ex}
\end{dblfig}

In this program, coordination assumes a far more important role than
we have seen before. Coordination rules drive the behavior of the
algorithm and while the result of the algorithm is statistically
identical to the original algorithm, SBP works very differently than
LBP.  SBP is also implemented in GraphLab~\cite{GraphLab2010}, a C++
framework for writing machine algorithms.  GraphLab provides the
splash scheduler as part of its the framework. It includes 350 lines
of C++ code.  With our coordination facts, it is possible to create
the necessary scheduling with only 12 rules.

### 2014-11-13 sssp.tex
space saving edits

Fig.~\ref{results:sssp_uspowergrid} shows experimental results when the
program computes the SSSP for 20\% of the nodes of the graph.
The scalability results show that the coordinated
version using 16 threads is 1.3 times faster than the regular version using 16
threads while still showing good scaling. Overall, the coordinated version produces 40\% less facts
derived when compared to the regular version.
Note that there are some situations where unnecessary facts are propagated
because although the shortest distance is selected locally, sub-optimal distances may be
propagated because many SSSP distances are computed at the same time.

>>>>

Fig.~\ref{results:sssp_uspowergrid} shows experimental results when
the program computes the SSSP for 20\% of the nodes of the graph.  The
coordinated version scales well and compared to the regular version
using 16 threads is 1.3 times faster.  the coordinated version
produces 40\% fewer facts when compared to the regular version.
Note that there are some situations where unnecessary facts are
propagated because although the shortest distance is selected locally,
sub-optimal distances may be propagated because many SSSP distances
are computed at the same time.
